--- 
title: "Ceramic rim morphology differs significantly between contexts at the Carson site"
author: "Robert Z. Selden, Jr."
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "Morphology of ceramic rims."
---

# Preface {-}

>Reproducibility---the ability to recompute results---and replicability---the chances other experimenters will achieve a consistent result---are two foundational characteristics of successful scientific research [@RN20918].

This volume is written using _Markdown_, and includes all analysis code employed in the study, providing a means for others to reproduce (exactly) those results discussed and expounded upon in the following chapters. The replicable nature of this undertaking provides a means for others to critically assess and evaluate the various analytical components of this study [@RN20915;@RN20916;@RN20917], which is a necessary requirement for the production of reliable knowledge.

## Basis of inquiry

Using the figures and tables provided in a recent article published in [_Southeastern Archaeology_](https://doi.org/10.1080/0734578X.2021.1922072), rim images and tabular data were repurposed for a study of traditional and geometric morphometrics to assess whether ceramic rim shape differs between two contexts at the Carson site.

## Primary findings

### Traditional morphometrics
* No variables differ significantly between contexts

### Geometric morphometrics
* Rim shape **differs significantly** by context

## Data management

The data and analysis code associated with this project can be accessed through this document or the GitHub repository [https://github.com/aksel-blaise/carson.rim](https://github.com/aksel-blaise/carson.rim), which is digitally curated on the Open Science Framework [DOI: 10.17605/OSF.IO/AE9ZM](https://osf.io/ae9zm/).

## Colophon

This version of the analysis was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# what R packages and versions were used?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

Current Git commit details are:

```{r}
# where can I find this commit? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
